---
title: "Currency Exchange Dashboard"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
    theme: readable
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(tibble)
library(forcats)
library(lubridate)
library(forecast)
library(prophet)
library(xgboost)
library(zoo)
library(lmtest)
library(igraph)
library(ggraph)

history <- read_csv("../data/history.csv") %>% drop_na()
recent <- read_csv("../data/recent.csv")

common_currencies <- c("USD", "KRW", "GBP", "JPY", "AUD", "CAD", "CHF")

exchange_data <- bind_rows(history, recent) %>%
  mutate(date = ymd(date)) %>%
  arrange(date)

log_returns <- exchange_data %>%
  filter(currency %in% common_currencies) %>%
  arrange(currency, date) %>%
  group_by(currency) %>%
  mutate(log_return = log(rate / lag(rate))) %>%
  ungroup() %>%
  select(date, currency, log_return) %>%
  pivot_wider(names_from = currency, values_from = log_return)

run_forecasts <- function(data, currency_code) {
  df <- data %>% 
    filter(currency == currency_code) %>%
    arrange(date) %>%
    mutate(
      date_num = as.numeric(date),
      weekday = wday(date),
      month = month(date)
    )

  # Create lag features for XGBoost
  for (i in 1:5) {
    df[[paste0("lag_", i)]] <- lag(df$rate, i)
  }
  df <- df %>% drop_na()

  # Train XGBoost model
  xgb_matrix <- xgb.DMatrix(
    data = as.matrix(df %>% select(starts_with("lag_"), date_num, weekday, month)),
    label = df$rate
  )
  model_xgb <- xgboost(data = xgb_matrix, nrounds = 100, objective = "reg:squarederror", verbose = 0)

  # Generate XGBoost forecasts
  future_dates <- seq(max(df$date) + 1, by = "day", length.out = 7)
  lags <- tail(df$rate, 5)
  preds <- numeric(7)

  for (i in 1:7) {
    input <- tibble(
      lag_1 = lags[5], lag_2 = lags[4], lag_3 = lags[3],
      lag_4 = lags[2], lag_5 = lags[1],
      date_num = as.numeric(future_dates[i]),
      weekday = wday(future_dates[i]),
      month = month(future_dates[i])
    )
    preds[i] <- predict(model_xgb, xgb.DMatrix(as.matrix(input)))
    lags <- c(lags[-1], preds[i])
  }

  xgb_df <- tibble(date = future_dates, predicted = preds)

  # ARIMA forecast
  ts_data <- ts(df$rate)
  arima_model <- auto.arima(ts_data)
  arima_fc <- forecast(arima_model, h = 7)
  arima_df <- tibble(date = future_dates, predicted = as.numeric(arima_fc$mean))

  # Prophet forecast
  prophet_df <- df %>% select(ds = date, y = rate)
  model_prophet <- prophet(prophet_df)
  future_df <- make_future_dataframe(model_prophet, periods = 7)
  prophet_fc <- predict(model_prophet, future_df)
  
  # Extract only the last 7-day forecast
  prophet_df_final <- prophet_fc %>%
    filter(ds > max(df$date)) %>%
    select(date = ds, predicted = yhat)
  
  return(list(
    xgboost = xgb_df,
    arima = arima_df,
    prophet = prophet_df_final,
    prophet_full = prophet_fc,
    raw = df
  ))
}

plot_visualization <- function(currency_code) {
  raw_data <- results[[currency_code]]$raw %>%
    arrange(date)

  max_point <- raw_data %>% filter(rate == max(rate, na.rm = TRUE)) %>% dplyr::slice(1)
  min_point <- raw_data %>% filter(rate == min(rate, na.rm = TRUE)) %>% dplyr::slice(1)

  ggplot(raw_data, aes(x = date, y = rate)) +
    geom_line(color = 'steelblue') +
    geom_point(data = max_point, aes(x = date, y = rate), color = "red", size = 3) +
    geom_point(data = min_point, aes(x = date, y = rate), color = "blue", size = 3) +
    geom_text(data = max_point, aes(label = paste("Max:", round(rate, 2))), vjust = -1, hjust = 0.5, size = 3) +
    geom_text(data = min_point, aes(label = paste("Min:", round(rate, 2))), vjust = 1.5, hjust = 0.5, size = 3) +
    labs(title = paste("Exchange Rate for", currency_code),
         x = "Date", y = "Rate")
}

plot_log_return <- function(currency_code) {
  raw_data <- results[[currency_code]]$raw %>%
    arrange(date) %>%
    mutate(log_return = log(rate / lag(rate))) %>%
    filter(!is.na(log_return))  # Remove NA values

  ggplot(raw_data, aes(x = date, y = log_return)) +
    geom_line(color = "darkred") +
    labs(
      title = paste("Log Return of", currency_code),
      x = "Date",
      y = "Log Return"
    )
}

plot_xgboost_forecast <- function(currency_code, recent_months = 1) {
  raw_data <- results[[currency_code]]$raw
  pred_data <- results[[currency_code]]$xgb
  ggplot() +
    geom_line(data = raw_data %>% filter(date > max(date) - months(recent_months)),
              aes(x = date, y = rate), color = "gray") +
    geom_line(data = pred_data, aes(x = date, y = predicted), color = "green") +
    labs(title = paste("XGBoost Forecast for", currency_code))
}

plot_arima_forecast <- function(currency_code, recent_days = 21) {
  raw_data <- results[[currency_code]]$raw
  pred_data <- results[[currency_code]]$arima
  ggplot() +
    geom_line(data = raw_data %>% filter(date > max(date) - days(recent_days)),
              aes(x = date, y = rate), color = "gray") +
    geom_line(data = pred_data, aes(x = date, y = predicted), color = "purple") +
    labs(title = paste("ARIMA Forecast for", currency_code))
}

plot_recent_prophet <- function(forecast_df, history_df, days_before = 21) {
  last_train_date <- max(history_df$date)

  # match time zone
  tzone <- attr(forecast_df$ds, "tzone")
  last_train_date <- as.POSIXct(last_train_date, tz = tzone)

  start_plot <- last_train_date - days(days_before)

  forecast_df %>%
    filter(ds >= start_plot) %>%
    ggplot(aes(x = ds, y = yhat)) +
    geom_ribbon(aes(ymin = yhat_lower, ymax = yhat_upper), fill = "lightblue", alpha = 0.3) +
    geom_line(color = "blue") +
    labs(title = "Prophet Forecast (Recent)", x = "Date", y = "Rate")
}

run_granger_analysis <- function(log_returns, currencies, max_lag = 3) {
  results <- expand.grid(from = currencies, to = currencies) %>%
    filter(from != to) %>%
    rowwise() %>%
    mutate(
      p_value = tryCatch({
        grangertest(log_returns[[to]] ~ log_returns[[from]], order = max_lag)$`Pr(>F)`[2]
      }, error = function(e) NA_real_)
    ) %>%
    ungroup() %>%
    arrange(p_value)

  return(results)
}

plot_granger_heatmap <- function(granger_results, currencies) {
  library(ggplot2)
  
  # Convert to matrix format for heatmap
  heatmap_data <- granger_results %>%
    complete(from = currencies, to = currencies, fill = list(p_value = NA)) %>%
    pivot_wider(names_from = to, values_from = p_value) %>%
    column_to_rownames("from") %>%
    as.matrix()
  
  # Reshape for ggplot
  heatmap_df <- as.data.frame(as.table(heatmap_data))
  colnames(heatmap_df) <- c("From", "To", "P_value")
  
  ggplot(heatmap_df, aes(x = To, y = From, fill = P_value)) +
    geom_tile(color = "white") +
    scale_fill_gradient(low = "red", high = "white", na.value = "grey90") +
    labs(
      title = "Granger Causality P-Value Heatmap",
      x = "To (Dependent Variable)",
      y = "From (Independent Variable)",
      fill = "p-value"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

plot_granger_network <- function(granger_results, alpha = 0.05) {
  # Filter significant causal relationships
  filtered <- granger_results %>% filter(!is.na(p_value), p_value < alpha)

  # Create directed graph
  g <- graph_from_data_frame(filtered, directed = TRUE)

  # Visualize graph
  ggraph(g, layout = "fr") +
    geom_edge_link(aes(label = round(p_value, 3)),
                   arrow = arrow(length = unit(3, 'mm')),
                   end_cap = circle(3, 'mm'),
                   label_size = 2, edge_width = 0.8) +
    geom_node_point(size = 5, color = "steelblue") +
    geom_node_text(aes(label = name), repel = TRUE) +
    labs(title = "Granger Causality Network (p < 0.05)") +
    theme_void()
}

summarize_granger_influence <- function(granger_results, alpha = 0.05) {
  granger_results %>%
    filter(p_value < alpha) %>%
    group_by(from) %>%
    summarise(Influences_Others = n(), .groups = "drop") %>%
    full_join(
      granger_results %>%
        filter(p_value < alpha) %>%
        group_by(to) %>%
        summarise(Influenced_By_Others = n(), .groups = "drop"),
      by = c("from" = "to")
    ) %>%
    rename(Currency = from) %>%
    replace_na(list(Influences_Others = 0, Influenced_By_Others = 0)) %>%
    arrange(desc(Influences_Others))
}

results <- list()
for (cur in common_currencies) {
  results[[cur]] <- run_forecasts(exchange_data, cur)
}
```

# Forecast Dashboard {.tabset}

## USD {.tabset}

### Visualization
```{r}
plot_visualization("USD")
plot_log_return("USD")
```

### ARIMA
```{r}
plot_arima_forecast("USD")
```

### Prophet
```{r}
plot_recent_prophet(results$USD$prophet_full, results$USD$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("USD")
```

## KRW {.tabset}

### Visualization
```{r}
plot_visualization("KRW")
plot_log_return("KRW")
```

### ARIMA
```{r}
plot_arima_forecast("KRW")
```

### Prophet
```{r}
plot_recent_prophet(results$KRW$prophet_full, results$KRW$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("KRW")
```

## GBP {.tabset}

### Visualization
```{r}
plot_visualization("GBP")
plot_log_return("GBP")
```

### ARIMA
```{r}
plot_arima_forecast("GBP")
```

### Prophet
```{r}
plot_recent_prophet(results$GBP$prophet_full, results$GBP$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("GBP")
```

## JPY {.tabset}

### Visualization
```{r}
plot_visualization("JPY")
plot_log_return("JPY")
```

### ARIMA
```{r}
plot_arima_forecast("JPY")
```

### Prophet
```{r}
plot_recent_prophet(results$JPY$prophet_full, results$JPY$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("JPY")
```

## AUD {.tabset}

### Visualization
```{r}
plot_visualization("AUD")
plot_log_return("AUD")
```

### ARIMA
```{r}
plot_arima_forecast("AUD")
```

### Prophet
```{r}
plot_recent_prophet(results$AUD$prophet_full, results$AUD$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("AUD")
```

## CAD {.tabset}

### Visualization
```{r}
plot_visualization("CAD")
plot_log_return("CAD")
```

### ARIMA
```{r}
plot_arima_forecast("CAD")
```

### Prophet
```{r}
plot_recent_prophet(results$CAD$prophet_full, results$CAD$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("CAD")
```

## CHF {.tabset}

### Visualization
```{r}
plot_visualization("CHF")
plot_log_return("CHF")
```

### ARIMA
```{r}
plot_arima_forecast("CHF")
```

### Prophet
```{r}
plot_recent_prophet(results$CHF$prophet_full, results$CHF$raw)
```

### XGBoost
```{r}
plot_xgboost_forecast("CHF")
```


# Granger Causality Analysis {.tabset}
```{r}
# Run Granger causality analysis after computing log returns
granger_results <- run_granger_analysis(log_returns, common_currencies, max_lag = 3)

# Display only statistically significant results
granger_results %>% filter(p_value < 0.05)

# Plot p-value heatmap
plot_granger_heatmap(granger_results, common_currencies)

plot_granger_network(granger_results)

influence_summary <- summarize_granger_influence(granger_results)
print(influence_summary)
```

